\singlespacing \vspace{-1.5cm} \LARGE Trade-offs among plant floral and reproductive traits determine interactions with floral visitors

### 2021-08-12 Draft/Manuscript.tex
Paper almost ready

\documentclass[12pt,a4paper,]{article}
\usepackage[]{mathpazo}
\usepackage{setspace}
\setstretch{2}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            pdftitle={Trade-offs among plant floral and reproductive traits determine interactions with floral visitors},
            pdfauthor={Jose B. Lanuza1,2*; Romina Rader1; Jamie Stavert3; Liam K. Kendall4; Manu E. Saunders1; Ignasi Bartomeus2},
            colorlinks=true,
            linkcolor=RoyalBlue,
            citecolor=Blue,
            urlcolor=RoyalBlue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

% Add Supplementary Tables and Figures
% Code from https://stackoverflow.com/a/51337664
\newcommand{\beginsupplement}{
  \setcounter{table}{0}  
  \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} 
  \renewcommand{\thefigure}{S\arabic{figure}}
}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{float}
\floatplacement{figure}{H}
\usepackage[skip=3pt]{caption}
\captionsetup[figure]{labelformat=empty}
\usepackage{setspace}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Trade-offs among plant floral and reproductive traits determine
interactions with floral visitors}
\author{Jose B. Lanuza\textsuperscript{1,2}* \and Romina Rader\textsuperscript{1} \and Jamie Stavert\textsuperscript{3} \and Liam K. Kendall\textsuperscript{4} \and Manu E. Saunders\textsuperscript{1} \and Ignasi Bartomeus\textsuperscript{2}}
\date{}

@@

\small
\doublespacing

\textsuperscript{1} School of Environmental and Rural Science,
University of New England, Armidale, New South Wales 2350, Australia

\textsuperscript{2} Estación Biológica de Doñana (EBD-CSIC), E-41092
Seville, Spain

\textsuperscript{3} Department of Conservation \textbar{} Te Papa
Atawhai, Auckland, New Zealand

\textsuperscript{4} Centre for Environmental and Climate Science, Lund
University, Sölvegatan 37, S-223 62 Lund, Sweden

\texttt{*} Corresponding author:
\href{mailto:barragansljose@gmail.com}{\nolinkurl{barragansljose@gmail.com}}

\normalsize
\singlespacing

@@

\doublespacing
\vspace{5mm} \bleft

There is an astonishing diversity of floral structures and plant
reproductive strategies among flowering plants (Barrett, 2002; Schiestl
\& Johnson, 2013), which have long been of interest to pollination
biologists in terms of their relevance to plant-pollinator interactions.
However, most studies that have explored reproductive (e.g., mating and
compatibility systems) and floral trait (e.g., flower size or nectar
provision) variation have concentrated on the individual or community
level and thus, broader macroecological patterns remain poorly
investigated (Carvalheiro et al., 2014; Baude et al., 2016; Munoz et
al., 2016; Moeller et al., 2017; Grossenbacher et al., 2017). Indeed,
studies depicting species' life history strategies generally focus on
vegetative traits and rarely consider reproductive traits
(Salguero-Gómez et al., 2016; Rüger et al., 2018). As a consequence, a
unified framework that explores the compromises among floral traits and
their relevance to plant life strategies is currently lacking (Roddy et
al., 2021). At the same time, there is growing interest in the
determinants of plant-pollinator interactions via trait-based approaches
(Fenster et al., 2004) and trait-matching analyses (Bartomeus et al.,
2016). However, floral traits have been overlooked beyond highly
specialised plant-pollinator systems (Dellinger, 2020; Roddy et al.,
2021) and the role of plant reproductive biology remains little explored
in plant-pollinator interactions (but see Tur et al., 2013 and Devaux et
al., 2014).

With the recent availability of large trait databases, plant ecological
strategies are increasingly being examined (e.g., TRY, Kattge et al.,
2011 and COMPADRE, Salguero‐Gómez et al., 2015), and are facilitating
the identification of global patterns and constraints of plant form and
function (Salguero‐Gómez et al., 2015; Díaz et al., 2016; Carmona et
al., 2021). However, the main focus has been on vegetative traits such
as leaf (Wright et al., 2004) or wood (Chave et al., 2009) trade-offs
with little or no attention given to reproductive and floral traits
(E‐Vojtkó et al., 2020), also critical to plant form and function. For
instance, short lived versus perennial species tend to have low versus
high levels of outcrossing, respectively (Barrett, 2003; Moeller et al.,
2017). Further, outcrossing levels are positively correlated with flower
size (Goodwillie et al., 2010). In addition, the presence of costly
rewards (e.g., pollen or nectar) and showy flowers or floral displays
can only be understood through consideration of plant species' reliance
upon animal pollination (pollinator dependence) and their role in
attracting pollinators (Ollerton et al., 2011). Hence, exploring plant
life strategies with reproductive and floral trade-offs, in conjunction
with their pollinator dependence, is necessary for a balanced
understanding of plant economics.

Several studies have identified links between plant traits and
plant-pollinator network properties (Lázaro et al., 2008; Bartomeus,
2013; Carvalheiro et al., 2014). Moreover, plant traits can also define
species' network roles (e.g., specialists vs generalists). For example,
species that occupy trait space extremes are more likely to exhibit
higher specialization and be more reliant on trait-matching (Junker et
al., 2013; Coux et al., 2016). This morphological matching between plant
and floral visitors can determine plant-pollinator interactions, and
thus shape their interaction network structure (Stang et al., 2009;
Ibanez, 2012). Despite the increasing knowledge of the relevance of
traits on the species network roles, little is known about how plant
reproductive and floral traits determine plant species' network roles at
a macroecological scale.

@@

\textbf{Plant strategies}

The phylogenetically informed principal component analysis (pPCA)
captured by the first two and three axes 51.8\% and 70.97\% of trait
variation, respectively (Figure 1 and Figure S5) and had a phylogenetic
correlation (\textbf{lambda}) of 0.76. The first principal component
(PC1) represented 26.72\% of the trait variation and indicated a
trade-off between flower number and flower size. We refer to this axis
as the `flower number - flower size trade-off as already described in
previous studies (Sargent et al., 2007; Kettle et al., 2011). Hence, one
end of the spectrum comprised species with high investment in flower
number and plant height but small flower size, short style length and
low ovule number. The other end of this spectrum comprised species that
were short in height and invested in large flowers, long styles, many
ovules, but few flowers. The main contributing traits to PC1 were plant
height, flower number, ovule number and flower size (loadings
\textgreater{} \textbar{}0.5\textbar{}; Table S3) but style length also
contributed moderately on PC1 (loading = -0.33). The second principal
component (PC2) represented 25.05\% of the trait variation and indicated
a trade-off between low and high pollinator dependence. We refer to this
axis as the `pollinator dependence trade-off'. The main driver of trait
variation on PC2 was autonomous selfing (loading = 0.85) but the other
traits (except ovule number) also made moderate contributions (loadings
from 0.27 to 0.4; Table S3). We found that high pollinator dependence
was associated with larger and higher numbers of flowers, greater plant
height and longer styles. In contrast, species with high levels of
autonomous selfing tended to have fewer and smaller flowers, had shorter
styles and were shorter in height. Further, PC3 explained a considerable
amount of trait variability (19.17\%) and the main contributors to this
axis were style length (loading = -0.66) and the degree of autonomous
selfing (loading = -0.51). The remaining traits, apart from ovule
number, were moderately correlated to changes on PC3 (loadings from
-0.23 to -0.46; Table S3). Thus, because style length was correlated
with all traits on PC3 and was the main driver of variation, we refer to
this axis as the `style length trade-off'. Further, the pPCA with the
subset of species that had nectar and pollen quantity data showed that
nectar quantity (microlitres of nectar per flower) was positively
associated with flower size, style length and ovule number (PC1,
23.40\%); and pollen quantity (pollen grains per flower) was positively
correlated with flower number and plant height and negatively associated
with autonomous selfing (PC2, 21.67\%; Figure S6). This pPCA explained
similar variance with the first two principal components (45.07\%) and
similar associations of traits despite some variability in the loadings
(Table S4).

@@

We found that most categorical traits were statistically associated with
the first two axes of trait variation (Figure 2 and Table S2). Flower
symmetry, which was only associated with PC2 (Sum of squares = 8.51,
F-value = 14.72, p \textless{} 0.01 ), and nectar provision, which was
independent of PC1 and PC2 (PC1: Sum of squares = 0.37, F-value = 0.29 ,
p = 0.59; PC2: Sum of squares = 0.83, F-value = 1.43, p = 0.23) showed
lack of statistical association. In addition, we found statistical
differences (Tukey test) between the different levels of categorical
traits in the trait space (Figure S7). Regarding self compatibility, we
found larger differences on PC2 (i.e., species with unisexual flowers
that were self incompatibile were statistically differentiated from
species with partial or full self compatibility; Figure S7 A and B).
Life forms differed statistically across both axes of trait variation
and followed a gradient of larger life forms (trees and shrubs) with
higher pollinator dependence to smaller ones (herbs) with lower
pollinator dependence (Figure S7 C and D). Consequently, lifespan also
followed this gradient but perennial and short lived species only
differed statistically on PC2 (Figure S7 E and F). Species with
unisexual flowers (monoecious and dioecious) were clustered on both
extremes of the first two principal components and had the highest
pollinator dependence and highest number of flowers (Figure S7 G and H).
Moreover, we found that the campanulate and capitulum flower shapes were
differentiated from tube, papilionaceous, open and brush shapes in the
trait space. The former morphologies had larger flowers and greater
pollinator dependence, while the latter had higher flower number and
greater autonomous selfing (Figure S7 I and J). Regarding flower
symmetry, zygomorphic flowers were associated with lower levels of
pollinator dependence, whereas actinomorphic flowers had higher levels
of pollinator dependence (Figure S7 K and L).

@@

@@

\begin{figure}[H]
\includegraphics{output/figures/unnamed-chunk-3-1} \caption{Location of the different qualitative traits that showed statistical association with the first two axes of trait variation. These different traits included: compatibility system (a), life form (b), lifespan (c), breeding system (d), flower shape (e) and flower symmetry (f).}\label{fig:unnamed-chunk-3}
\end{figure}

@@

\textbf{Phylogenetic signal of traits}

\subsubsection{OLD}\label{old}

Among flowering plants there is an astonishing diversity of floral
structures and plant reproductive strategies (Barrett
\protect\hyperlink{ref-barrett2002}{2002}, Schiestl and Johnson
\protect\hyperlink{ref-schiestl2013}{2013}). Subsequently, pollination
biologists have long studied their relevance on plant-pollinator
interactions. However, despite recent efforts exploring reproductive
(e.g., mating and compatibility system) and floral traits (e.g., flower
size or nectar provision) with large sets of species (Carvalheiro et al.
\protect\hyperlink{ref-carvalheiro2014}{2014}, Baude et al.
\protect\hyperlink{ref-baude2016}{2016}, Munoz et al.
\protect\hyperlink{ref-munoz2016}{2016}, Grossenbacher et al.
\protect\hyperlink{ref-grossenbacher2017}{2017}, Moeller et al.
\protect\hyperlink{ref-moeller2017}{2017}), most studies focus on the
individual, community level or specific taxa and macroecological
patterns remain poorly investigated. For instance, studies depicting
species' life history strategies generally focus on vegetative traits
and rarely consider reproductive traits on the main axes of trait
variation (Díaz et al. \protect\hyperlink{ref-diaz2016}{2016},
Salguero-Gómez et al. \protect\hyperlink{ref-salguero2016}{2016}). In
addition, there is not a unified framework that explores the compromises
of floral traits and their relevance on plant life strategies (Roddy et
al. \protect\hyperlink{ref-roddy2021}{2021}). Similarly, there is
growing interest in understanding what determines plant-pollinator
network structure. Approaches with traits range from the description of
pollination syndromes (e.g., Dellinger
\protect\hyperlink{ref-dellinger2020}{2020}), to more specific
trait-matching analysis (Bartomeus et al.
\protect\hyperlink{ref-bartomeus2016}{2016}), but again the reproductive
biology of the species has received little attention (but see Tur et al.
\protect\hyperlink{ref-tur2013}{2013} and @devaux2014) and floral traits
have been overlooked beyond highly specialised plant-pollinator
interactions (Dellinger \protect\hyperlink{ref-dellinger2020}{2020},
Roddy et al. \protect\hyperlink{ref-roddy2021}{2021}).

>>>>

\documentclass[12pt,a4paper,]{article}
\usepackage[]{mathpazo}
\usepackage{setspace}
\setstretch{2}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            pdftitle={ Trade-offs among plant floral and reproductive traits determine interactions with floral visitors},
            colorlinks=true,
            linkcolor=RoyalBlue,
            citecolor=Blue,
            urlcolor=RoyalBlue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

% Add Supplementary Tables and Figures
% Code from https://stackoverflow.com/a/51337664
\newcommand{\beginsupplement}{
  \setcounter{table}{0}  
  \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} 
  \renewcommand{\thefigure}{S\arabic{figure}}
}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{float}
\floatplacement{figure}{H}
\usepackage[skip=3pt]{caption}
\captionsetup[figure]{labelformat=empty}
\usepackage{setspace}
\usepackage{titlesec}
\titlespacing{\title}{0pt}{\parskip}{-\parskip}
\usepackage[style=nature]{biblatex}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{\singlespacing \vspace{-1.5cm} \LARGE Trade-offs among plant floral and
reproductive traits determine interactions with floral visitors}
\author{}
\date{\vspace{-2.5em}}

@@

\vspace{-2.2cm} \singlespacing
\large 
\textbf{Jose B. Lanuza$^{1,2}*$, Romina Rader$^{1}$, Jamie Stavert$^{3}$, Liam K. Kendall$^{4}$, Manu E. Saunders$^{1}$ and Ignasi Bartomeus$^{2}$}

\normalsize

@@

\small
\vspace{-0.5cm}

\noindent\rule{\textwidth}{1pt}

\textsuperscript{1} School of Environmental and Rural Science,
University of New England, Armidale, New South Wales 2350, Australia.
\textsuperscript{2} Estación Biológica de Doñana (EBD-CSIC), E-41092
Seville, Spain. \textsuperscript{3} Department of Conservation
\textbar{} Te Papa Atawhai, Auckland, New Zealand. \textsuperscript{4}
Centre for Environmental and Climate Science, Lund University,
Sölvegatan 37, S-223 62 Lund, Sweden. \texttt{*} e-mail:
\href{mailto:barragansljose@gmail.com}{\nolinkurl{barragansljose@gmail.com}}

\doublespacing
\vspace{5mm} \bleft
\normalsize

There is an astonishing diversity of floral structures and plant
reproductive strategies among flowering
plants\textsuperscript{\protect\hyperlink{ref-barrett2002}{1},\protect\hyperlink{ref-schiestl2013}{2}},
which have long been of interest to pollination biologists in terms of
their relevance to plant-pollinator interactions. However, most studies
that have explored reproductive (e.g., mating and compatibility systems)
and floral trait (e.g., flower size or nectar provision) variation have
concentrated on the individual or community level and thus, broader
macroecological patterns remain poorly
investigated\textsuperscript{\protect\hyperlink{ref-carvalheiro2014}{3}--\protect\hyperlink{ref-grossenbacher2017}{7}}.
Indeed, studies depicting species' life history strategies generally
focus on vegetative traits and rarely consider reproductive
traits\textsuperscript{\protect\hyperlink{ref-salguero2016}{8}}. As a
consequence, a unified framework that explores the compromises among
floral traits and their relevance to plant life strategies is currently
lacking\textsuperscript{\protect\hyperlink{ref-roddy2021}{10}}. At the
same time, there is growing interest in the determinants of
plant-pollinator interactions via trait-based
approaches\textsuperscript{\protect\hyperlink{ref-fenster2004}{11}} and
trait-matching
analyses\textsuperscript{\protect\hyperlink{ref-bartomeus2016}{12}}.
However, floral traits have been overlooked beyond highly specialised
plant-pollinator
systems\textsuperscript{\protect\hyperlink{ref-roddy2021}{10},\protect\hyperlink{ref-dellinger2020}{13}}
and the role of plant reproductive biology remains little explored in
plant-pollinator interactions (but see
references\textsuperscript{\protect\hyperlink{ref-tur2013}{14},\protect\hyperlink{ref-devaux2014}{15}}).

With the recent availability of large trait databases, plant ecological
strategies are increasingly being
examined\textsuperscript{\protect\hyperlink{ref-kattge2011}{16},\protect\hyperlink{ref-salguero2015}{17}},
and are facilitating the identification of global patterns and
constraints of plant form and
function\textsuperscript{\protect\hyperlink{ref-salguero2016}{8},\protect\hyperlink{ref-diaz2016}{18},\protect\hyperlink{ref-carmona2021}{19}}.
However, the main focus has been on vegetative traits such as
leaf\textsuperscript{\protect\hyperlink{ref-wright2004}{20}} or
wood\textsuperscript{\protect\hyperlink{ref-chave2009}{21}} trade-offs
with little or no attention given to reproductive and floral
traits\textsuperscript{\protect\hyperlink{ref-evojtko2020}{22}}, also
critical to plant form and function. For instance, short lived versus
perennial species tend to have low versus high levels of outcrossing,
respectively\textsuperscript{\protect\hyperlink{ref-moeller2017}{6},\protect\hyperlink{ref-barrett2003}{23}}.
Further, outcrossing levels are positively correlated with flower
size\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{24}}. In
addition, the presence of costly rewards (e.g., pollen or nectar) and
showy flowers or floral displays can only be understood through
consideration of plant species' reliance upon animal pollination
(pollinator dependence) and their role in attracting
pollinators\textsuperscript{\protect\hyperlink{ref-ollerton2011}{25}}.
Hence, exploring plant life strategies with reproductive and floral
trade-offs, in conjunction with their pollinator dependence, is
necessary for a balanced understanding of plant economics.

Several studies have identified links between plant traits and
plant-pollinator network
properties\textsuperscript{\protect\hyperlink{ref-carvalheiro2014}{3},\protect\hyperlink{ref-lazaro2008}{26},\protect\hyperlink{ref-bartomeus2013}{27}}.
Moreover, plant traits can also define species' network roles (e.g.,
specialists vs generalists). For example, species that occupy trait
space extremes are more likely to exhibit higher specialization and be
more reliant on
trait-matching\textsuperscript{\protect\hyperlink{ref-junker2013}{28},\protect\hyperlink{ref-coux2016}{29}}.
This morphological matching between plant and floral visitors can
determine plant-pollinator interactions, and thus shape their
interaction network
structure\textsuperscript{\protect\hyperlink{ref-stang2009}{30},\protect\hyperlink{ref-ibanez2012}{31}}.
Despite the increasing knowledge of the relevance of traits on the
species network roles, little is known about how plant reproductive and
floral traits determine plant species' network roles at a
macroecological scale.

@@

\textbf{Plant strategies.} The phylogenetically informed principal
component analysis (pPCA) captured by the first two and three axes
51.8\% and 70.97\% of trait variation, respectively (Fig. 1 and
Supplementary Fig. S5) and had a phylogenetic correlation (\(\lambda\))
of 0.76. The first principal component (PC1) represented 26.72\% of the
trait variation and indicated a trade-off between flower number and
flower size. We refer to this axis as the `flower number - flower size
trade-off as already described in previous
studies\textsuperscript{\protect\hyperlink{ref-sargent2007}{32},\protect\hyperlink{ref-kettle2011}{33}}.
Hence, one end of the spectrum comprised species with high investment in
flower number and plant height but small flower size, short style length
and low ovule number. The other end of this spectrum comprised species
that were short in height and invested in large flowers, long styles,
many ovules, but few flowers. The main contributing traits to PC1 were
plant height, flower number, ovule number and flower size (loadings
\textgreater{} \textbar{}0.5\textbar{}; Supplementary Table S3) but
style length also contributed moderately on PC1 (loading = -0.33). The
second principal component (PC2) represented 25.05\% of the trait
variation and indicated a trade-off between low and high pollinator
dependence. We refer to this axis as the `pollinator dependence
trade-off'. The main driver of trait variation on PC2 was autonomous
selfing (loading = 0.85) but the other traits (except ovule number) also
made moderate contributions (loadings from 0.27 to 0.4; Supplementary
Table S3). We found that high pollinator dependence was associated with
larger and higher numbers of flowers, greater plant height and longer
styles. In contrast, species with high levels of autonomous selfing
tended to have fewer and smaller flowers, had shorter styles and were
shorter in height. Further, PC3 explained a considerable amount of trait
variability (19.17\%) and the main contributors to this axis were style
length (loading = -0.66) and the degree of autonomous selfing (loading =
-0.51). The remaining traits, apart from ovule number, were moderately
correlated to changes on PC3 (loadings from -0.23 to -0.46;
Supplementary Table S3). Thus, because style length was correlated with
all traits on PC3 and was the main driver of variation, we refer to this
axis as the `style length trade-off'. Further, the pPCA with the subset
of species that had nectar and pollen quantity data showed that nectar
quantity (microlitres of nectar per flower) was positively associated
with flower size, style length and ovule number (PC1, 23.40\%); and
pollen quantity (pollen grains per flower) was positively correlated
with flower number and plant height and negatively associated with
autonomous selfing (PC2, 21.67\%; Supplementary Fig. S6). This pPCA
explained similar variance with the first two principal components
(45.07\%) and similar associations of traits despite some variability in
the loadings (Supplementary Table S4).

@@

We found that most categorical traits were statistically associated with
the first two axes of trait variation (Fig. 2 and Supplementary Table
S2). Flower symmetry, which was only associated with PC2 (Sum of squares
= 8.51, F-value = 14.72, p \textless{} 0.01 ), and nectar provision,
which was independent of PC1 and PC2 (PC1: Sum of squares = 0.37,
F-value = 0.29 , p = 0.59; PC2: Sum of squares = 0.83, F-value = 1.43, p
= 0.23) showed lack of statistical association. In addition, we found
statistical differences (Tukey test) between the different levels of
categorical traits in the trait space (Supplementary Fig. S7). Regarding
self compatibility, we found larger differences on PC2 (i.e., species
with unisexual flowers that were self incompatibile were statistically
differentiated from species with partial or full self compatibility;
Supplementary Fig. S7 A and B). Life forms differed statistically across
both axes of trait variation and followed a gradient of larger life
forms (trees and shrubs) with higher pollinator dependence to smaller
ones (herbs) with lower pollinator dependence (Supplementary Fig. S7 C
and D). Consequently, lifespan also followed this gradient but perennial
and short lived species only differed statistically on PC2
(Supplementary Fig. S7 E and F). Species with unisexual flowers
(monoecious and dioecious) were clustered on both extremes of the first
two principal components and had the highest pollinator dependence and
highest number of flowers (Supplementary Fig. S7 G and H). Moreover, we
found that the campanulate and capitulum flower shapes were
differentiated from tube, papilionaceous, open and brush shapes in the
trait space. The former morphologies had larger flowers and greater
pollinator dependence, while the latter had higher flower number and
greater autonomous selfing (Supplementary Fig. S7 I and J). Regarding
flower symmetry, zygomorphic flowers were associated with lower levels
of pollinator dependence, whereas actinomorphic flowers had higher
levels of pollinator dependence (Supplementary Fig. S7 K and L).

@@

\textbf{Phylogenetic signal of traits.} We found a strong phylogenetic
signal (P \textless{} 0.01) in all quantitative traits (Supplementary
Table S5). The traits that showed the highest phylogenetic signal were
ovule number (\(\lambda = 1\)), pollen grains per flower
(\(\lambda = 1\)) and plant height (\(\lambda = 0.96\)), followed by
flower length (\(\lambda = 0.75\)), flower width (\(\lambda = 0.73\)),
number of flowers per plant (\(\lambda = 0.69\)) and nectar
concentration (\(\lambda = 0.65\)). The traits that showed a moderate
phylogenetic signal were inflorescence width (\(\lambda = 0.57\)), style
length (\(\lambda = 0.49\)) and autonomous selfing (\(\lambda = 0.34\)).
Finally, microliters of nectar per flower showed the lowest phylogenetic
signal of all traits (\(\lambda = 0.14\)).

\textbf{Visitation patterns.} The main axes of trait variation explained
little of the overall visitation rates (conditional \(R2 = 0.31\);
marginal \(R2 = 0.06\)) but showed relevant trends when we explored the
interaction with the different floral visitors guilds (Fig. 3). All
floral visitors guilds visited plant species with higher pollinator
dependence more frequently (PC2; Fig. 3b and Fig. 3e). In addition, all
guilds other than Syrphids and Lepidoptera (i.e., all Hymenoptera,
non-Syrphidae-Diptera and Coleoptera) showed greater visitation rates on
species with small numerous flowers (PC1; Fig. 3a and Fig. 3d). On the
style length trade-off (PC3), Anthophila-Hymenoptera, Lepidoptera and
Non-Anthophila-Hymenoptera showed greater visitation rates on plant
species with larger styles and higher levels of selfing; while
Syrphidae, Non-Syrphidae-Diptera and Coleoptera showed higher visitation
rates on species with shorter styles and lower selfing (Fig. 3c and Fig.
3f). Furthermore, when running an additional model that separates the
most represented families of Anthophila-Hymenoptera (bees;
\(marginal R2 = 0.30\); \(conditional R2 = 0.03\)) showed that the
family Apidae was the main driver of the observed patterns
(Supplementary Fig. S8).

@@

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}\textbf{Fig. 2 \textbar{} Fitted
posterior estimates of the number of visits made by the different floral
visitors guilds in relation to the three main axes of trait variation
(PC1, PC2 and PC3)}. Bees (a, b and c) and the rest of guilds (d, e and
f) were plotted separately for visualization purposes. In addition, we
trimmed the plotting area that was over the 95th percentile to improve
visualisation. PC1 represents the flower number - flower size trade-off,
PC2 represents the pollinator dependence trade-off and PC3, the style
length trade-off.}
\end{figure}

@@

\textbf{Plant species functional roles.} The three main axes of trait
variation poorly explain the variance of the different plant species
level metrics (Supplementary Fig. S9; interaction frequency
\textasciitilde{} PCs, \(conditional R2 = 0.11\),
\(marginal R2 = 0.02\); normalized degree \textasciitilde{} PCs,
\(conditional R2 = 0.24\), \(marginal R2 = 0.02\); and, specialization
\textasciitilde{} PCs, \(conditional R2 = 0.37\),
\(marginal R2 = 0.03\)). Overall, the most notable trends were found on
PC1 and PC3 for interaction frequency and specialization. On the flower
number - flower size trade-off (PC1), interaction frequency was higher
for plant species with more flowers but was lower for plant species with
larger flowers. On PC1, specialization showed the opposite trend. On the
style length trade-off (PC3), interaction frequency was lower for plants
with shorter styles and lower selfing and higher for species with long
styles and higher levels of selfing. Again, specialization showed the
opposite trend to interaction frequency.

When we further investigate which combination of traits drive plant
network roles, we show that the regression tree for visitation frequency
was best explained by plant height, nectar concentration and style
length (Fig. 4a). Specifically, species taller than 3.9 m had the
highest interaction frequency, while species that were shorter than 3.9
m and had a nectar concentration lower than 16\% had the lowest
interaction frequency. Normalized degree was best explained by nectar
concentration, pollen grains per flower, plant height, flower width and
autonomous selfing (Fig. 4b). Species with a nectar concentration over
49\% had the highest levels of normalized degree, whereas species with
nectar concentration lower than 49\%, more than 21,000 pollen grains per
flower and height less than 0.78 m had the lowest normalized degree.
Finally, specialization was best explained by plant height, ovule
number, pollen grains per flower and autonomous selfing (Fig. 4c).
Overall, plant species with the highest specialization were shorter than
1.3 m, had more than 14,000 pollen grains per flower and autonomously
self-pollinated less than 11\% of their fruits. In contrast, species
taller or equal than 5.1 m and with lower than 14 ovules per flower had
the lowest specialization values.

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-4-1.pdf}
\caption{\label{fig:unnamed-chunk-4}\textbf{Fig. 4 \textbar{} Regression
tree analysis of interaction frequency (log-transformed), normalized
degree and specialization for the subset of species with quantitative
data for pollen and nectar traits}. The superior value inside the node
indicates the mean value of the different species-level metric and the
lower value, the percentage of species that are considered in each node.
Thus, the top node has the mean value of the named trait for the 100\%
of species. Each node has a yes/no question and when the condition is
fulfilled, the branch turns to the `yes' direction and when not, to the
`no' direction. This rationale is followed in all the regression trees
as indicated in the first branch division of the topmost node of each
tree.}
\end{figure}
